---
title: "UBUG Presentation"
author: "Craig Demke"
date: "`r Sys.Date()`"
output: html_document
---

```{r load libraries, include=FALSE}
##library(gt)
library(here)
##library(shiny)
##library(purrr)
##library(emo)
##library(DT)
library(knitr)
library(dplyr)
```

# Welcome 🎉   
- Who am I
 - masteR manipulatoR 
- What is R
 - Language
 - Code
 - Functions
 - Packages
 - usheUtils 

# Language:  Syntax vs. Semantics
```{r eval = TRUE, echo = TRUE}
syntax <- "sentence rules and grammar"
semantics <- "meaning"

print(syntax)
print(semantics)
```

class: center
## Appreciation for Language
learning another language helps you with your native language.
```{r eval = TRUE, echo = TRUE}
appreciate <- "sette pris på" 
word_to_ask <- "appriciate"
it_broken_down <- "to set a price on"

paste("hvordan sier man", word_to_ask, "på norsk?", sep = " ")
print(appreciate)
print(it_broken_down)
```
R is like learning a new language as an adult after having already learned your native tongue.

# Code: 
Latin 'codex', which means "a book of laws"
- R is an open source scripting language for predictive analytics and data visualization.
- RStudio is an integrated development environment (IDE) for R and Python.
- I want you to think about the first code you learned?  

## Unfiltered: 
"Hey rodent, help save the Galaxy"

## Coded politely:
**Pardon me** *Rocket*, **I'm sorry** to bother you I know you're busy, I was wondering if you could **please** help me save the Galaxy.

## Making Two List in R
```{r eval = TRUE, echo = TRUE}
social_interactions <- c("I need something", 
"I may need it again later",
"youre in my way",
"I may need something from you in the future",
"pay attention to me",
"undo",
"possible incoming request",
"important variable")
```

```{r eval = TRUE, echo = TRUE}
polite_code <- c("Please", 
"Thank you",
"Pardon me",
"You're welcome",
"Excuse me",
"I'm sorry",
"May I help you",
"person's name")
```

## Making a data frame
```{r eval = TRUE, echo = TRUE}
polite_df <- data.frame(social_interactions, polite_code) 
knitr::kable(polite_df)
```

## Manipulating Data 
So let's say I want to remember the two things I need when asking for something. 
I could filter the dataframe for only those variables.
```{r eval = TRUE, echo = TRUE}
polite_code_for_favor <- polite_df %>% # think of this %>% symbol as an "and"
  dplyr::filter(social_interactions == "I need something" 
                | social_interactions == "important variable")

knitr::kable(polite_code_for_favor)
```

### Arguments?
In the Console
```{r eval = FALSE, echo = TRUE}
?dplyr::filter()
```

# Functions: 
A function in R is an object containing multiple interrelated statements that are run together in a predefined order every time the function is called.

## Function Example
```{r eval = TRUE, echo = TRUE}
ask_for_help <- function(persons_name, request) {
paste("Pardon me", 
      persons_name, 
      "I was wondering if you could help me", 
      request,
      "Thanks.", 
      sep = " ")
}
```

### Now we use the function
```{r eval = TRUE, echo = TRUE}
test <- ask_for_help("Craig", "with an R problem.")
test2 <- ask_for_help("Randy", "with Argos")
print(test)
print(test2)
```

### Making a list of the Guardians of the Galaxy
```{r eval = TRUE, echo = TRUE}
gog_names <- c("Star Lord", 
"Gamorrah,",
"Drax,",
"Rocket,",
"Groot,",
"Nebula,",
"Mantis,",
"James Gunn's brother,")
```

### Making a list of offences
```{r eval = TRUE, echo = TRUE}
offences <- c("a trash panda.",
              "a walking thesaurus.",
              "this dumb tree.",
              "the biggest idot in the Galaxy.",
               "an imbecile.",
               "ugly.",
               "I am Groot.",
               "a losser.")
```

## Making a new Data Frame
```{r eval = TRUE, echo = TRUE}
offended_df <- data.frame(
  offender = sample(gog_names, 8, replace = FALSE),
  offence = sample(offences, 8, replace = FALSE),
  offended = sample(gog_names, 8, replace = TRUE)
  )
```

```{r eval = TRUE, echo = TRUE}
knitr::kable(offended_df)
```

## Apology Function 
```{r eval = TRUE, echo = TRUE}
apologize <- function(input_df) {
  output_df <- input_df %>% 
    dplyr::mutate(apology = paste("Hey", 
           offended,
           "this is",
           offender,
           "sorry I called you,",
           '"',
           offence,
           '"',
           sep = " ")) %>%
     dplyr::select(apology)

return(output_df)
}
```

Here we can pass the apologize function
and use the whole data frame as the only argument
```{r eval = TRUE, echo = TRUE}
random_apologies <- apologize(offended_df)
```

## Apology Function 
```{r eval = TRUE, echo = TRUE}
random_apologies <- apologize(offended_df)
knitr::kable(random_apologies)
```

# Packages:
R packages are a collection of R functions, complied code and sample data.

## Package Example:
- The apologize function could be part of an etiquette package
- install.packages("etiquette")  
 - etiquette::apologize()
 - etiquette::ask_for_help() 
 - etiquette::respond_to_request()
 - etiquette::move_pass_someone()
 - etiquette::repair_offense()
 - etiquette::lower_your_volume()
 - etiquette::reciprocity_for_future_request()
 - etiquette::insert_humor_in_presentation()

# the Package usheUtils:: 
A R package created by Utah Tech which contains functions for most of the USHE
reports. There are functions for each part of a report including meta-functions that
run all the functions for the report as a whole. https://github.com/dsu-effectiveness/usheUtils

You could simply install usheUtils from Utah Tech's Github repo from inside R studio 
```{r eval = FALSE, echo = TRUE}
install.packages("devtools")
devtools::install_github("ut-effectiveness/usheUtils")
```
Then run your instution's data tables through the correct function 
```{r eval = FALSE, echo = TRUE}
usheUtils::generate_student_course_submission_file()
usheUtils::generate_dws_submission_file()
usheUtils::generate_financial_aid_submission_file()
usheUtils::generate_course_submission_file()
usheUtils::generate_graduation_submission_file()
usheUtils::generate_program_submission_file()
usheUtils::generate_space_inventory_building_submission_file()
usheUtils::generate_space_inventory_rooms_submission_file()
usheUtils::generate_student_submission_file()
```
---